|

### 2021-07-19 paper/paper.tex

\author[FirstAuthorLastName \textit{et~al}.]{
Stephen D. Turner\,\textsuperscript{1}, V.P.
Nagraj\,\textsuperscript{1}, Matthew
Scholz\,\textsuperscript{1}, Shakeel Jessa\,\textsuperscript{1}, Carlos
Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2,3}, August
Woener\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
}

@@

The skater package has several functions for importing, parsing, and
analyzing pedigree data. Pedigrees define familial relationships in a
hierarchical structure. Many genomics tools for working with pedigrees
start with a .fam file, which is a tabular format with one row per
individual and columns for unique IDs of the mother, father, and the
family unit. The skater package contains the function
\texttt{read\_fam()} to read in a PLINK-formatted .fam file and another
function \texttt{fam2ped()} to convert the content into a pedigree
object as a nested tibble with one row per family. All pedigree
processing from skater internally leverages a data structure from the
kinship2 package \citep{kinship2}. Further functions such as
\texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of
the pedigree for each family, while \texttt{ped2kinpair()} produces a
pairwise list of relationships between all individuals in the data with
the expected kinship coefficients for each pair.

@@

Once estimated kinship is converted to degree, it may be of interest to
compare the inferred degree to known degrees of relatedness. When
aggregated over many relationships and inferences, this can help
benchmark performance of a particular kinship analysis method. The
skater package adapts a \texttt{confusion\_matrix()} function from
\citet{clark2021} to provide standard contingency table metrics
(e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a
new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more
thoroughly in the skater package vignette (see Supplementary Material)
and may be a preferable measure of classification accuracy when
benchmarking relationship degree estimation. Taking the reciprocal of
the target and predicted degree in a typical RMSE calculation results in
larger penalties for more egregious misclassifications (e.g.,
classifying a first-degree relative pair as second-degree) than
misclassifications at more distant relationships (e.g., classifying a
fourth-degree relative pair as fifth-degree).

>>>>

\author[FirstAuthorLastName \textit{et~al}.]{
Stephen D. Turner\,\textsuperscript{1}, V.P.
Nagraj\,\textsuperscript{1}, Matthew
Scholz\,\textsuperscript{1}, Shakeel Jessa\,\textsuperscript{1}, Carlos
Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2,3}, August
E. Woerner\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
}

@@

The skater package has several functions for importing, parsing, and
analyzing pedigree data. Pedigrees define familial relationships in a
hierarchical structure. Many genomics tools for working with pedigrees
start with a .fam file, which is a tabular format with one row per
individual and columns for unique IDs of the mother, father, and the
family unit. The skater package contains the function
\texttt{read\_fam()} to read in a PLINK-formatted .fam file and another
function \texttt{fam2ped()} to convert the content into a pedigree
object as a nested tibble with one row per family. All pedigree
processing from skater internally leverages a data structure from the
kinship2 package \citep{kinship2}. Further functions such as
\texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of
the pedigree for each family, while \texttt{ped2kinpair()} produces a
pairwise list of relationships between all individuals in the data with
the expected kinship coefficients for each pair (see Supplementary
Material).

@@

Once estimated kinship is converted to degree, it may be of interest to
compare the inferred degree to known degrees of relatedness. When
aggregated over many relationships and inferences, this can help
benchmark performance of a particular kinship analysis method. The
skater package adapts a \texttt{confusion\_matrix()} function from
\citet{clark2021} to provide standard contingency table metrics
(e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a
new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more
thoroughly in the skater package vignette (see Supplementary Material)
and may be a preferable measure of classification accuracy when
benchmarking relationship degree estimation. In many kinship
benchmarking analyses, classification error is treated in a categorical
manner (exact match plus or minus one degree), neglecting the true
amount of sharing as a real number. Taking the reciprocal of the target
and predicted degree in a typical RMSE calculation results in larger
penalties for more egregious misclassifications (e.g., classifying a
first-degree relative pair as second-degree) than misclassifications at
more distant relationships (e.g., classifying a fourth-degree relative
pair as fifth-degree).

### 2021-07-19 paper/supplement-vignette.Rmd

One of the formats used by PLINK and other genetic analysis tools is the `.fam` file,[^plink-fam] which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit. The package includes `read_fam()` to read files in this format:

@@

Family structures imported from "fam" formated files can then be translated to the `pedigree` structure used by the `kinship2` package.[^kinship2-vignette] The "fam" format may include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:

@@

The `plot_pedigree()` function from `skater` will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from each family:

>>>>

One of the formats used by PLINK and other genetic analysis tools is the `.fam` file.[^plink-fam] A `.fam` file is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit. The package includes `read_fam()` to read files in this format:

@@

Family structures imported from ".fam" formated files can then be translated to the `pedigree` structure used by the `kinship2` package.[^kinship2-vignette] The "fam" format may include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:

@@

The `plot_pedigree()` function from `skater` will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from family:
